---
title: "DiffBind HCA"
author: Rylee K. Hackley
output: html_notebook
---
```{r, message=FALSE}
# BiocManager::install(c("rtracklayer", "DiffBind"))
library(tidyverse)
library(rtracklayer)
library(DiffBind)
```

fragment sizes calculated by CHIPQC
WT_IP_0glu_A        104
WT_IP_0glu_B		    213
trmB_IP_0glu_A	    104
trmB_IP_0.1glu_A    167
trmB_IP_0glu_B	    200
trmB_IP_0.1glu_B	  103
trmB_IP_0glu_C      103
trmB_IP_0glu_D	    204	

```{r, message=F, warning=F, fig.height=6, fig.width=6}
# load all tagged samples and greylist
(meta <- read_csv("01b_HCA_chip_meta.csv")[3:8, ])
config <- list(
  minQCth = 30, fragmentSize = c(104, 167, 200, 103, 103, 204),
  bCorPlot = T, th = 0.05
)
grey <- import.bed("01b_greyList/WT_greyList.bed")

# create initial DBA object
trmB <- dba(sampleSheet = meta, peakFormat = "bed", minOverlap = 0.75, config = config)

# number of overlapping peaks in samples considered together or split by condition
dba.overlap(trmB, mode = DBA_OLAP_RATE)
dba.overlap(trmB, trmB$masks$`0.1_glucose`, mode = DBA_OLAP_RATE)
dba.overlap(trmB, trmB$masks$no_glucose, mode = DBA_OLAP_RATE)
```
total of 439 unique peaks across all samples.  
137 are shared between both replicates with glucose. 
140 are shared between at least 3 replicates with no glucose.

Mask peaks identified in WT control samples
```{r, fig.height=6, fig.width=6}
trmB <- dba.blacklist(trmB, blacklist = FALSE, greylist = grey)
dba.overlap(trmB, mode = DBA_OLAP_RATE)
dba.overlap(trmB, trmB$masks$`0.1_glucose`, mode = DBA_OLAP_RATE)
dba.overlap(trmB, trmB$masks$no_glucose, mode = DBA_OLAP_RATE)
```
total of 432 unique peaks across all samples.  
121 are shared between both replicates with glucose. 
124 are shared between at least 3 replicates with no glucose.

Calculate a binding matrix with scores based on read counts for every sample (affinity scores), rather than confidence scores for only those peaks called in a specific sample (occupancy scores).
```{r}
trmB <- dba.count(trmB, summits = 150, minOverlap = 0.75) # peaks must occur in 75% of replicates to be considered consensus peaks
trmB
```
100 peaks in the consensus, greylisted peakset

Normalize, set up contrasts, and execute differential binding analysis
```{r}
trmB <- dba.normalize(trmB, normalize = DBA_NORM_RLE, library = DBA_LIBSIZE_PEAKREADS) # DBA norm RLE has the best baseline, and identifies the most shared sites.
trmB <- dba.contrast(trmB, reorderMeta = list(Condition = "0.1_glucose"), minMembers = 2)
trmB <- dba.analyze(trmB)
dba.plotMA(trmB, method = DBA_DESEQ2)
dba.show(trmB, bContrasts = TRUE)
```

Visualize DBA results
```{r}
dba.plotHeatmap(trmB, contrast = 1)

# PCA for all sites in affinity matrix
dba.plotPCA(trmB, DBA_CONDITION, label = DBA_CONDITION)

# PCA plot using affinity data for only differential bound sites
dba.plotPCA(trmB, contrast = 1, label = DBA_CONDITION)

# binding affinity heatmap showing affinities for differential bound sites
hmap <- colorRampPalette(c("red", "black", "green"))(n = 13)
dba.plotHeatmap(trmB,
  contrast = 1,
  correlations = FALSE, scale = "row", colScheme = hmap
)
```

view report, compile score, and export
```{r}
(trmB.DB <- dba.report(trmB, bFlip = T, bCalled = T, bCounts = T, bNormalized = F))
trmB.DB$score <- trmB.DB$Fold
export.bed(trmB.DB, "02_DiffBind/02_diffbind2022.bed")
```
