---
title: "Comparing growth of TrmB deletion strains"
author: Rylee K. Hackley
output: 
  html_document:
    keep_md: true
---
```{r, message=FALSE, warning=FALSE, echo=FALSE}
# load required libraries
library(tidyverse)
library(viridis)
library(growthcurver)
library(scales)
library(ggpubr)
library(rstatix)

# custom function, calculated the 95% Confidence interval
conf_int95 <- function(data) {
  n <- length(data)
  error <- qt(0.975, df = n - 1) * sd(data) / sqrt(n)
  return(error)
}
```

Load the normalized data and meta data.
```{r}
expt <- "trmB_deletion_compare"

# set path to files
data <- "raw_growth_data/20220227_hca_trmBs_data.csv"
meta <- "raw_growth_data/20220227_hca_trmBs_meta.csv"

## read in files
mt <- read.csv(meta, stringsAsFactors = F)
mt$expt <- rep(expt, nrow(mt))
dt <- read.csv(data)

# set well no as column names
colnames(dt) <- c("time", seq(1, ncol(dt) - 1))

# remove the first 3 hours, or 6 timepoints.
dt0 <- dt
dt0 <- dt0[-(1:6), ]
dt0$time <- seq(0, length.out = nrow(dt0), by = 0.5) # redo time points
dt0 <- dt0[complete.cases(dt0), ]
```

```{r}
unite(mt, "ID", c("strain", "condition"), sep = "+")$ID %>%
  str_replace_all(., "\\+$", "") -> ID

# Save new metafile as a different variable. This is good practice!
mt0 <- cbind("variable" = mt$well, ID, mt[-1], stringsAsFactors = F)
mt0 <- mt0 %>% arrange(variable)

## remove rows based on observation (input own row names!!)
mt0 <- mt0 %>% dplyr::slice(-c(10, 110, 20, 88))
dt0 <- dt0 %>% select(!c(`10`, `110`, `20`, `88`))
```

```{r, collapse=TRUE}
# Convert data from wide to long format
m_dt <- reshape2::melt(dt0, id = "time")
mtdt <- merge(m_dt, mt0, by = "variable")
mtdt$environment <- paste(mtdt$media, mtdt$condition)
well <- paste(mtdt$variable, mtdt$ID, sep = " ")
mtdt <- cbind(well, mtdt)

# time cutoff - this is a useful function for graphing and analysis. You might not need to look at the full length of the experiment.
t_mtdt <- mtdt[!(mtdt$time >= 80), ]

unique(t_mtdt$condition)
unique(t_mtdt$strain)
```

remove blank outliers and plot.
```{r}
cond <- "25mM glucose"

filter(t_mtdt, strain == "blank" & condition == cond) %>%
  ggplot(., aes(x = time, y = value, color = well)) + # if you want to know the well number, change color = well
  ggtitle(paste("Growthcurves in", expt)) +
  xlab("time (h)") +
  ylab("Absorbance at 600nm") +
  geom_line() +
  scale_color_viridis(discrete = TRUE) +
  theme_classic() +
  theme( # legend.position = 'none',
    plot.title = element_text(face = "bold", color = "#000000", size = 12, angle = 0),
    axis.title.x = element_text(face = "bold", color = "#000000", size = 10, angle = 0),
    axis.title.y = element_text(face = "bold", color = "#000000", size = 10, angle = 90),
    axis.text.y = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    axis.text.x = element_text(face = "plain", color = "#000000", size = 10, angle = 0)
  )

filter(t_mtdt, strain == "DF60" & condition == cond) %>%
  ggplot(., aes(x = time, y = value, color = well)) + # if you want to know the well number, change color = well
  ggtitle(paste("Growthcurves in", expt)) +
  xlab("time (h)") +
  ylab("Absorbance at 600nm") +
  geom_line() +
  scale_color_viridis(discrete = TRUE) +
  theme_classic() +
  theme( # legend.position = 'none',
    plot.title = element_text(face = "bold", color = "#000000", size = 12, angle = 0),
    axis.title.x = element_text(face = "bold", color = "#000000", size = 10, angle = 0),
    axis.title.y = element_text(face = "bold", color = "#000000", size = 10, angle = 90),
    axis.text.y = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    axis.text.x = element_text(face = "plain", color = "#000000", size = 10, angle = 0)
  )

filter(t_mtdt, strain == "AKS133" & condition == cond) %>%
  ggplot(., aes(x = time, y = value, color = well)) + # if you want to know the well number, change color = well
  ggtitle(paste("Growthcurves in", expt)) +
  xlab("time (h)") +
  ylab("Absorbance at 600nm") +
  geom_line() +
  scale_color_viridis(discrete = TRUE) +
  theme_classic() +
  theme( # legend.position = 'none',
    plot.title = element_text(face = "bold", color = "#000000", size = 12, angle = 0),
    axis.title.x = element_text(face = "bold", color = "#000000", size = 10, angle = 0),
    axis.title.y = element_text(face = "bold", color = "#000000", size = 10, angle = 90),
    axis.text.y = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    axis.text.x = element_text(face = "plain", color = "#000000", size = 10, angle = 0)
  )

filter(t_mtdt, strain == "AKS319" & condition == cond) %>%
  ggplot(., aes(x = time, y = value, color = well)) + # if you want to know the well number, change color = well
  ggtitle(paste("Growthcurves in", expt)) +
  xlab("time (h)") +
  ylab("Absorbance at 600nm") +
  geom_line() +
  scale_color_viridis(discrete = TRUE) +
  theme_classic() +
  theme( # legend.position = 'none',
    plot.title = element_text(face = "bold", color = "#000000", size = 12, angle = 0),
    axis.title.x = element_text(face = "bold", color = "#000000", size = 10, angle = 0),
    axis.title.y = element_text(face = "bold", color = "#000000", size = 10, angle = 90),
    axis.text.y = element_text(face = "plain", color = "#000000", size = 10, angle = 0),
    axis.text.x = element_text(face = "plain", color = "#000000", size = 10, angle = 0)
  )
```
difference btwn ATCC33960 bioreps (a grows to higher stat than b in glucose, no carb)

average blank wells
```{r}
filter(mt0, strain == "blank")$variable -> blanks
rowMeans(dt0[colnames(dt0) %in% blanks]) -> blank
dt <- cbind("blank" = blank, dt0[, !names(dt0) %in% blanks])

# subtract average blank from each well
d <- dt[-c(1, 2)] - dt$blank
d$time <- dt$time # add time column back
mins <- Rfast::colMins(as.matrix(d), value = T) # get column minimums
d <- sweep(d, 2, mins, "-") # subtract smallest value for each column
```

Growth curve modeling
```{r}
# Create an output data frame to store the results in.
num_analyses <- length(names(d)) - 1
d_gc <- data.frame(
  sample = character(num_analyses),
  k = numeric(num_analyses),
  n0 = numeric(num_analyses),
  r = numeric(num_analyses),
  t_mid = numeric(num_analyses),
  t_gen = numeric(num_analyses),
  auc_l = numeric(num_analyses),
  auc_e = numeric(num_analyses),
  sigma = numeric(num_analyses),
  stringsAsFactors = FALSE
)

# Truncate or trim the input data to observations occurring in the first 80 hours.
trim_at_time <- 80.5

# Loop through all of the columns in the data frame. For each column, run Growthcurver.
pdf(paste("figures/", expt, "_growthcurver_fits.pdf", sep = ""), height = 8.5, width = 11)
par(mfcol = c(8, 12))
par(mar = c(0.25, 0.25, 0.25, 0.25))
y_lim_max <- max(d[, setdiff(names(d), "time")], na.rm = T) - min(d[, setdiff(names(d), "time")], na.rm = T)

n <- 1 # keeps track of the current row in the output data frame
for (col_name in names(d)) {
  # Don't process the time column.
  if (col_name != "time") {
    # Create a temporary data frame that contains just the time and current col
    d_loop <- d[c("time", col_name)]

    # Now, call Growthcurver to calculate the metrics using SummarizeGrowth
    gc_fit <- SummarizeGrowth(
      data_t = d_loop$time,
      data_n = d_loop[col_name],
      t_trim = trim_at_time,
      bg_correct = "none"
    )

    if (gc_fit$vals[[16]] == "cannot fit data") {
      d_gc$sample[n] <- col_name
      n <- n + 1
    } else {
      d_gc$sample[n] <- col_name
      d_gc[n, 2:9] <- c(
        gc_fit$vals$k,
        gc_fit$vals$n0,
        gc_fit$vals$r,
        gc_fit$vals$t_mid,
        gc_fit$vals$t_gen,
        gc_fit$vals$auc_l,
        gc_fit$vals$auc_e,
        gc_fit$vals$sigma
      )

      n <- n + 1
      # Finally, plot the raw data and the fitted curve. print some of the data points to keep the file size smaller
      n_obs <- length(gc_fit$data$t)
      idx_to_plot <- 1:20 / 20 * n_obs
      plot(gc_fit$data$t[idx_to_plot], gc_fit$data$N[idx_to_plot],
        pch = 20,
        xlim = c(0, trim_at_time),
        ylim = c(0, y_lim_max),
        cex = 0.6, xaxt = "n", yaxt = "n"
      )
      text(x = trim_at_time / 4, y = y_lim_max, labels = col_name, pos = 1)
      lines(gc_fit$data$t, predict(gc_fit$model), col = "red")
    }
  }
}
dev.off()
```

graph fitted curves:
```{r, message=F}
fitted_curves <- list()

n <- 1 # keeps track of the current row in the output data frame
for (col_name in names(d)) {
  if (col_name != "time") {
    d_loop <- d[c("time", col_name)]
    gc_fit <- SummarizeGrowth(
      data_t = d_loop$time,
      data_n = d_loop[col_name],
      t_trim = trim_at_time,
      bg_correct = "none"
    )
    if (gc_fit$vals[[16]] == "cannot fit data") {
      d_gc$sample[n] <- col_name
      n <- n + 1
    } else {
      fitted_curves[[n]] <- predict(gc_fit$model)
      names(fitted_curves)[n] <- col_name
      n <- n + 1
    }
  }
}

tmp <- Filter(length, fitted_curves)
fits <- as.data.frame(tmp, col.names = names(tmp))
colnames(fits) <- names(tmp)
fits <- cbind("time" = d$time[d$time < trim_at_time], fits)

# re-zero lowest value
mins <- Rfast::colMins(as.matrix(fits), value = T) # get column minimums
fits <- sweep(fits, 2, mins, "-") # subtract smallest value for each column
```

```{r, message=FALSE, warning=FALSE}
# Convert data from wide to long format
m_dt <- reshape2::melt(fits, id = "time")
mtdt <- merge(m_dt, mt0, by = "variable")
well <- paste(mtdt$variable, mtdt$ID, sep = " ")
mtdt <- cbind(well, mtdt)

t_mtdt2 <- mtdt[!(mtdt$ID == ""), ]

stats2 <- t_mtdt2 %>%
  group_by(ID, strain, condition, time) %>%
  summarise(
    reps = length(value),
    average = mean(value),
    CI95 = conf_int95(value)
  )
unique(stats2$strain)

stats2$average <- stats2$average + 1

# all conditions tested.
stats2 %>%
  ggplot(., aes(x = time, y = log10(average), color = strain)) +
  xlab("time (h)") +
  ylab("log10(Optical Denisty)") +
  geom_line(linewidth = 1) +
  scale_x_continuous(limits = c(0, 72)) +
  geom_ribbon(aes(ymin = log10(average - CI95), ymax = log10(average + CI95), fill = strain, color = NULL), alpha = 0.2) +
  scale_fill_viridis("strain", discrete = TRUE, begin = 0.1, end = 0.9) +
  scale_color_viridis("strain", discrete = TRUE, begin = 0.1, end = 0.9) +
  theme_bw() +
  facet_wrap(~condition) +
  theme(
    axis.title.x = element_text(face = "bold", size = 10, angle = 0), axis.title.y = element_text(face = "bold", size = 10, angle = 90),
    axis.text.y = element_text(size = 8, angle = 0), axis.text.x = element_text(size = 8, angle = 0),
    axis.ticks.length.y.left = unit(.05, "cm"), strip.text = element_text(size = 10),
    strip.background = element_blank(), strip.placement = "outside"
  ) -> fitted.data
plot(fitted.data)

stats3 <- t_mtdt2 %>%
  group_by(ID, strain, condition, time, expt) %>%
  summarise(
    reps = length(value),
    average = mean(value) + 1,
    CI95 = conf_int95(value)
  )

stats3 %>%
  filter(condition == "no carb" | condition == "25mM glucose") %>%
  filter(strain != "AKS318" & strain != "ATCC33960") %>%
  ggplot(., aes(x = time, y = log10(average), color = strain)) +
  xlab("time (h)") +
  ylab("log10(Optical Denisty)") +
  geom_line(size = 1) +
  scale_x_continuous(limits = c(0, 80)) +
  geom_ribbon(aes(ymin = log10(average - CI95), ymax = log10(average + CI95), fill = strain, color = NULL), alpha = 0.2) +
  scale_fill_viridis("strain", discrete = TRUE, begin = 0.1, end = 0.9) +
  scale_color_viridis("strain", discrete = TRUE, begin = 0.1, end = 0.9) +
  theme_bw() +
  facet_wrap(~condition) +
  theme(
    axis.title.x = element_text(face = "bold", size = 12),
    axis.title.y = element_text(face = "bold", size = 12, angle = 90),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    legend.position = "none",
    axis.ticks.length.y.left = unit(.05, "cm"),
    strip.text = element_text(size = 12),
    strip.background = element_blank(),
    strip.placement = "outside"
  ) -> fitted.data
plot(fitted.data)

pdf(paste("figures/", expt, "_fitcurves.pdf", sep = ""), height = 3, width = 4, bg = "transparent")
plot(fitted.data)
dev.off()
```

#TrmB normalized to WT (by condition)
normalize trmB to WT. take average of WT bioreps. calculate ratios of individual trmB bioreps. then average and calc SD/std err
```{r, message=FALSE, warning=FALSE}
colnames(d_gc)[1] <- "variable"
d_gc[d_gc == 0] <- NA
d_gc <- d_gc[complete.cases(d_gc), ]
merge(d_gc, mt0) -> tmp

# t-test considering technical replicates distinct.
tmp %>%
  filter(condition != "25mM glycerol") %>%
  filter(strain != "AKS318" & strain != "ATCC33960") -> tmp2

vals <- tmp2 %>%
  group_by(condition) %>%
  t_test(auc_e ~ strain, comparisons = list(c("DF60", "AKS319"), c("DF60", "AKS133"), c("AKS319", "AKS133")), p.adjust.method = "fdr") %>%
  add_y_position(fun = "mean_sd")

ggboxplot(tmp2,
  x = "strain", y = "auc_e",
  fill = "strain", add = "jitter",
  facet.by = "condition", size = 0.8, ylab = "AUC"
) +
  stat_pvalue_manual(vals, label = "p.adj.signif", tip.length = 0.01) +
  scale_fill_viridis("x", discrete = TRUE, begin = 0.1, end = 0.9, alpha = 0.8) +
  ylim(0, 13) +
  theme_bw() +
  theme(
    axis.title.y = element_text(face = "bold", size = 12, angle = 90),
    axis.text.y = element_text(size = 10, angle = 0),
    axis.text.x = element_text(size = 10, angle = 0),
    axis.title.x = element_text(face = "bold", size = 12),
    axis.ticks.length.y.left = unit(.05, "cm"),
    strip.text = element_text(size = 12),
    strip.background = element_blank(),
    strip.placement = "outside",
    legend.position = "none"
  ) -> auc
plot(auc)

pdf(paste("figures/", expt, "_auc.pdf", sep = ""), height = 3, width = 4, bg = "transparent")
plot(auc)
dev.off()
```
